---
title: "experiments"
format: html
---

```{r}
library(tidyverse)
library(logolink)
```

```{r}
model_path <- file.path("netlogo", "roads", "Segregation Utility.nlogox")
model_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"
```

```{r}
setup_file <- create_experiment(
  name = "Venue Segragation Model Analysis",
  repetitions = 3,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 20,
  metrics = c(
  'index-exposure',
  'index-dissimilarity',
  'index-clustering',
  'index-centrality',
  'index-concentration'
  ),
  constants = list(
    "density-percent" = c(70),
    "%-similar-wanted" = 30,
    "number-of-ethnicities" = 2,
    "%-different-wanted" = 0,
    "%-proportion-majority" = 50,
    "neighborhood-distance" = c(2),
    "venues-number" = c(5),
    "venue-catchment-distance" = c(1, 3, 10, 15),
    "venue-exclusivity" = 1,
    "utility-threshold" = 0.1,
    "beta1-neighbors" = c(0.5),
    "beta2-venue-group" = 0.5,
    "beta3-venue-loc" = 0,
    "beta4-home-venues" = 0,
    "road-scenario" = 'empty',
    "parks-number" = 3,
    "park-catchment-distance" = 2,
    "beta5-park-group" = 0.25,
    "travel-distance" = 5,
    "road-multiplier" = 1,
    "real-travel?" = F,
    "random-seed?" = F
  )
)


```

```{r}
results <-
  model_path |>
  run_experiment(
    setup_file = setup_file,
    output = c("table")
  )
```

```{r}
library(magrittr)
data <-
  results |>
  extract2("table") 
```


### Segregation Indices Over Time by Neighborhood Distance

```{r}
data |> 
  filter(step <= 100) |>  
  select(
    run_number,
    step,
    starts_with("index_"),
    venue_catchment_distance         
  ) |> 
  pivot_longer(
    cols = starts_with("index_"),
    names_to = "index",
    values_to = "value"
  ) |> 
  filter(index %in% c("index_exposure", "index_dissimilarity", 
                      "index_clustering")) |>
  filter(venue_catchment_distance %in% c(1, 15)) |>
  mutate(grouping_combo_run = str_c(run_number, index, venue_catchment_distance),
         grouping_combo_condition = str_c(index, venue_catchment_distance)) |> 
  ggplot(aes(x = step, y = value, color = venue_catchment_distance |> factor(), group = grouping_combo_run, linetype = index)) +
  geom_line(alpha = 0.25)+
  # facet_wrap(~ venue_catchment_distance)+
  geomtextpath::geom_textpath(
    aes(group = index, label = index), 
    offset = unit(5, "pt"), gap = T, text_smoothing = 90)+
  labs(title = "Segregation Indices Over Time (venue_catchment_distance)", x = "Step", 
       y = "Index Value", color = "venue_catchment_distance") +
  theme_minimal()+
  theme(legend.position = "bottom")
       
```



```{r}
create_experiment_template(model_path, print = TRUE)

setup_file <- create_experiment(
  name = "Auto-Generated Analysis",
  repetitions = 1,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 200,
  metrics = c(
    "percent-similar",
    "percent-unhappy",
    "index-exposure",
    "index-dissimilarity",
    "index-clustering",
    "index-centrality",
    "index-concentration",
    "[pxcor] of patches",
    "[pycor] of patches",
    "[pcolor] of patches"
  ),
  constants = list(
    "density-percent" = c(71.0),
    "%-similar-wanted" = c(30.0),
    "number-of-ethnicities" = c(2.0),
    "%-different-wanted" = c(0.0),
    "neighborhood-distance" = c(1.0),
    "venues-number" = c(4.0),
    "venue-catchment-distance" = c(11.0),
    "venue-exclusivity" = c(1.0),
    "utility-threshold" = c(0.1),
    "beta1-neighbors" = c(0.1),
    "beta2-venue-group" = c(0.1),
    "beta3-venue-loc" = c(0.1),
    "beta4-home-venues" = c(0.1),
    "parks-number" = c(1.0),
    "park-catchment-distance" = c(10.0),
    "beta5-park-group" = c(0.1),
    "travel-distance" = c(9.0),
    "road-multiplier" = c(1.0),
    "real-travel?" = c(FALSE),
    "random-seed?" = c(FALSE),
    "new-venue-color" = c(15.0),
    "visual-mode" = c("abstract"),
    "road-scenario" = c("empty")
  )
)
```

# function 


```{r}
library(tidyverse)
library(logolink)
library(geomtextpath)

run_segregation_experiment <- function(
  model_path,
  repetitions = 3,
  max_steps = 100,
  indices_to_plot = c("index_exposure", "index_dissimilarity", "index_clustering"),
  # Pass constants as a list so you can easily modify them per run
  constants = list(
    "density-percent" = 70,
    "%-similar-wanted" = 30,
    "number-of-ethnicities" = 2,
    "%-different-wanted" = 0,
    "proportion-majority" = 50,
    "neighborhood-distance" = 2,
    "venues-number" = 5,
    "venue-catchment-distance" = 7, # Example of a tested/varying parameter
    "venue-exclusivity" = 1,
    "utility-threshold" = 0.1,
    "beta1-neighbors" = 0.5,
    "beta2-venue-group" = 0.5,
    "beta3-venue-loc" = 0,
    "beta4-home-venues" = 0,
    "road-scenario" = "empty",
    "parks-number" = 3,
    "park-catchment-distance" = 2,
    "beta5-park-group" = 0.25,
    "travel-distance" = 5,
    "road-multiplier" = 1,
    "real-travel?" = FALSE,
    "random-seed?" = FALSE
  )
) {
  
  # 1. Define the experiment setup
  setup_file <- create_experiment(
    name = "Venue Segregation Model Analysis",
    repetitions = repetitions,
    run_metrics_every_step = TRUE,
    setup = "setup",
    go = "go",
    time_limit = max_steps + 50, # Buffer to ensure we hit max_steps safely
    metrics = c(
      'index-exposure',
      'index-dissimilarity',
      'index-clustering',
      'index-centrality',
      'index-concentration'
    ),
    constants = constants
  )
  
  # 2. Run the experiment
  results <- model_path |>
    run_experiment(
      setup_file = setup_file,
      output = c("table")
    )
  
  # 3. Extract the table natively
  data <- results[["table"]]
  
  # 4. Identify the varying parameter to dynamically color the plot
  # Finds any constant with length > 1 (e.g., c(1, 15))
  varying_params <- names(constants)[lengths(constants) > 1]
  
  if (length(varying_params) == 0) {
    legend_title <- "run_number"
    varying_col <- "run_number"
  } else {
    legend_title <- varying_params[1]
    # NetLogo hyphenated params become underscored cols in R
    varying_col <- str_replace_all(legend_title, "-", "_") 
  }
  
  # 5. Build dynamic title string from all constants
  param_strings <- imap_chr(constants, ~ paste0(.y, ": ", paste(.x, collapse = ", ")))
  dynamic_subtitle <- str_wrap(paste(param_strings, collapse = " | "), width = 110)
  
  # 6. Process data and generate the plot
  p <- data |> 
    filter(step <= max_steps) |> 
    select(
      run_number,
      step,
      starts_with("index_"),
      all_of(varying_col) 
    ) |> 
    pivot_longer(
      cols = starts_with("index_"),
      names_to = "index",
      values_to = "value"
    ) |> 
    filter(index %in% indices_to_plot) |>
    mutate(
      # Tidy evaluation: Use .data[[varying_col]] to reference the column dynamically
      grouping_combo_run = str_c(run_number, index, .data[[varying_col]]),
      # Clean up the index names for prettier labels
      index_label = str_remove(index, "index_") 
    ) |> 
    ggplot(aes(
      x = step, 
      y = value, 
      color = factor(.data[[varying_col]]), 
      group = grouping_combo_run, 
      linetype = index_label
    )) +
    geom_line(alpha = 0.25) +
    geomtextpath::geom_textpath(
      aes(group = index_label, label = index_label), 
      offset = unit(5, "pt"), 
      gap = TRUE, 
      text_smoothing = 95,
      color = "gray20"
    ) +
    labs(
      title = "Segregation Indices Over Time",
      subtitle = varying_col, # Show the name of the varying parameter in the subtitle
      caption  = dynamic_subtitle,
      x = "Step", 
      y = "Index Value", 
      color = legend_title,
      linetype = "Index"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      plot.subtitle = element_text(size = 11, color = "gray20", margin = margin(b = 10)),
      plot.caption = element_text(size = 7, color = "gray50", hjust = 0, margin = margin(t = 15))
    )+
    guides(linetype = "none")
  
  # Return both the raw data and the plot as a list so you can access either
  return(list(
    data = data,
    plot = p
  ))
}


library(tidyverse)
library(logolink)
library(geomtextpath)



plot_segregation_results <- function(
  experiment_output,
  indices_to_plot = c("index_exposure", "index_dissimilarity", "index_clustering")
) {
  
  # Extract components from the experiment output
  data <- experiment_output$data
  constants <- experiment_output$constants
  max_steps <- experiment_output$max_steps
  
  # Identify the varying parameter to dynamically color the plot
  varying_params <- names(constants)[lengths(constants) > 1]
  
  if (length(varying_params) == 0) {
    legend_title <- "run_number"
    varying_col <- "run_number"
  } else {
    legend_title <- varying_params[1]
    varying_col <- str_replace_all(legend_title, "-", "_") 
  }
  
  # Build dynamic title string from all constants for the caption
  param_strings <- imap_chr(constants, ~ paste0(.y, ": ", paste(.x, collapse = ", ")))
  dynamic_subtitle <- str_wrap(paste(param_strings, collapse = " | "), width = 110)
  
  # Process data and generate the plot
  p <- data |> 
    filter(step <= max_steps) |> 
    select(
      run_number,
      step,
      starts_with("index_"),
      all_of(varying_col) 
    ) |> 
    pivot_longer(
      cols = starts_with("index_"),
      names_to = "index",
      values_to = "value"
    ) |> 
    filter(index %in% indices_to_plot) |>
    mutate(
      grouping_combo_run = str_c(run_number, index, .data[[varying_col]]),
      index_label = str_remove(index, "index_") 
    ) |> 
    ggplot(aes(
      x = step, 
      y = value, 
      color = factor(.data[[varying_col]]), 
      group = grouping_combo_run, 
    )) +
    geom_line(
      aes(linetype = index_label),
      alpha = 0.25,
      linewidth = 2) +
    # geomtextpath::geom_textpath(
    #   aes(group = index_label, label = index_label),
    #   offset = unit(5, "pt"),
    #   gap = T,
    #   text_smoothing = 95,
    #   color = "gray20"
    # ) +
    labs(
      title = "Segregation Indices Over Time",
      subtitle = varying_col, 
      caption  = dynamic_subtitle,
      x = "Step", 
      y = "Index Value", 
      color = legend_title,
      linetype = "Index"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      plot.subtitle = element_text(size = 11, color = "gray20", margin = margin(b = 10)),
      plot.caption = element_text(size = 7, color = "gray50", hjust = 0, margin = margin(t = 15))
    )
    # guides(linetype = "none")
  
  return(p)
}

library(tidyverse)
library(logolink)
library(geomtextpath)
library(colorspace) # Required for lightening the venue colors

run_segregation_model <- function(
  model_path,
  repetitions = 3,
  max_steps = 100,
  capture_spatial = FALSE, 
  constants = list(
    "density-percent" = 70,
    "%-similar-wanted" = 30,
    "number-of-ethnicities" = 2,
    "%-different-wanted" = 0,
    "proportion-majority" = 50,
    "neighborhood-distance" = 2,
    "venues-number" = 5,
    "venue-catchment-distance" = 7, 
    "venue-exclusivity" = 1,
    "utility-threshold" = 0.1,
    "beta1-neighbors" = 0.5,
    "beta2-venue-group" = 0.5,
    "beta3-venue-loc" = 0,
    "beta4-home-venues" = 0,
    "road-scenario" = "empty",
    "parks-number" = 3,
    "park-catchment-distance" = 2,
    "beta5-park-group" = 0.25,
    "travel-distance" = 5,
    "road-multiplier" = 1,
    "real-travel?" = FALSE,
    "random-seed?" = FALSE
  )
) {
  
  base_metrics <- c(
    'index-exposure',
    'index-dissimilarity',
    'index-clustering',
    'index-centrality',
    'index-concentration'
  )
  
  spatial_metrics <- c(
    '[pxcor] of patches',
    '[pycor] of patches',
    '[pcolor] of patches',
    '[is-road?] of patches',
    '[is-river?] of patches',
    '[xcor] of families',
    '[ycor] of families',
    '[color] of families',
    '[xcor] of venues',
    '[ycor] of venues',
    '[color] of venues'
  )
  
  metrics_to_run <- if (capture_spatial) c(base_metrics, spatial_metrics) else base_metrics
  outputs_to_run <- if (capture_spatial) c("table", "lists") else c("table")
  
  setup_file <- create_experiment(
    name = "Venue Segregation Model Analysis",
    repetitions = repetitions,
    run_metrics_every_step = TRUE,
    setup = "setup",
    go = "go",
    time_limit = max_steps + 50, 
    metrics = metrics_to_run,
    constants = constants
  )
  
  results <- model_path |> run_experiment(setup_file = setup_file, output = outputs_to_run)
  
  # -- NEW: Memory Optimization --
  # Identify constants that do NOT change (length == 1) to remove from outputs
  static_params <- names(constants)[lengths(constants) == 1]
  # Convert NetLogo parameter syntax to R column syntax
  static_cols <- str_replace_all(static_params, "-", "_")
  
  # Clean up the table output
  data_out <- results[["table"]]
  if (!is.null(data_out)) {
    cols_to_remove <- intersect(static_cols, names(data_out))
    data_out <- data_out |> select(-all_of(cols_to_remove))
  }
  
  # Clean up the lists output
  lists_out <- if (capture_spatial) results[["lists"]] else NULL
  if (!is.null(lists_out)) {
    cols_to_remove <- intersect(static_cols, names(lists_out))
    lists_out <- lists_out |> select(-all_of(cols_to_remove))
  }
  
  return(list(
    data = data_out,
    lists = lists_out,
    constants = constants,
    max_steps = max_steps
  ))
}


plot_segregation_world <- function(experiment_output, step_to_plot = 100, run_to_plot = 1) {
  
  if (is.null(experiment_output$lists)) {
    stop("Spatial data is missing. You must run `run_segregation_model()` with `capture_spatial = TRUE`.")
  }
  
  list_data <- experiment_output$lists |>
    filter(step == step_to_plot, run_number == run_to_plot)
  
  if (nrow(list_data) == 0) {
    stop("No data found for step ", step_to_plot, " in run ", run_to_plot)
  }

  # 1. Extract, clean, and parse patches safely
  patches <- list_data |>
    select(run_number, step, ends_with("_of_patches")) |>
    unnest(cols = ends_with("_of_patches")) |>
    rename_with(~ str_remove(.x, "_of_patches")) |>
    rename_with(~ str_replace_all(.x, "[\\?\\-]", "_")) |> 
    mutate(
      pcolor = parse_netlogo_color(as.numeric(pcolor)),
      pxcor = as.numeric(pxcor),
      pycor = as.numeric(pycor)
    )

  # Safely handle missing/empty road and river data
  if (!"is_road_" %in% names(patches)) patches$is_road_ <- FALSE
  if (!"is_river_" %in% names(patches)) patches$is_river_ <- FALSE

  # Apply patch color overrides
  patches <- patches |>
    mutate(
      is_road_ = as.logical(is_road_),
      is_river_ = as.logical(is_river_),
      render_color = case_when(
        is_river_ ~ "#4682B4", # Steel Blue
        is_road_ ~ "#333333",  # Dark Gray
        TRUE ~ pcolor
      )
    )

  # 2. Extract and parse families
  families <- list_data |>
    select(run_number, step, ends_with("_of_families")) |>
    unnest(cols = ends_with("_of_families")) |>
    rename_with(~ str_remove(.x, "_of_families")) |>
    mutate(
      color = parse_netlogo_color(as.numeric(color)),
      xcor = as.numeric(xcor),
      ycor = as.numeric(ycor)
    )

  # 3. Extract venues and check if they exist
  venue_cols <- names(list_data)[str_detect(names(list_data), "_of_venues$")]
  has_venues <- FALSE
  
  if (length(venue_cols) > 0) {
    venues <- list_data |>
      select(run_number, step, all_of(venue_cols)) |>
      unnest(cols = all_of(venue_cols))
    
    # Only process if unnesting resulted in actual rows
    if (nrow(venues) > 0) {
      has_venues <- TRUE
      venues <- venues |>
        rename_with(~ str_remove(.x, "_of_venues")) |>
        mutate(
          base_color = parse_netlogo_color(as.numeric(color)),
          fill_color = colorspace::lighten(base_color, amount = 0.3),
          xcor = as.numeric(xcor),
          ycor = as.numeric(ycor)
        )
    }
  }

  # 4. Generate the baseline plot (Environment + Families)
  p <- ggplot() +
    geom_tile(data = patches, aes(x = pxcor, y = pycor, fill = render_color), alpha = 0.7) +
    geom_point(data = families, aes(x = xcor, y = ycor, color = color), shape = 15, size = 1.9) +
    scale_fill_identity() + 
    scale_color_identity() + 
    coord_equal() +
    theme_void() +
    labs(
      title = "NetLogo World State",
      subtitle = paste("Run:", run_to_plot, "| Step:", step_to_plot)
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, color = "gray40")
    )

  # 5. Conditionally overlay Venues if they exist
  if (has_venues) {
    p <- p + geom_point(
      data = venues,
      aes(x = xcor, y = ycor, fill = fill_color),
      shape = 24,      
      color = "white", 
      size = 3.5,      
      stroke = 1     
    )
  }

  return(p)
}

library(patchwork) # Required for side-by-side ggplot arrangement

compare_segregation_runs <- function(
  experiment_output, 
  runs_to_compare = c(1, 2, 3), 
  step_to_plot = 100,
  ncol = NULL # Let patchwork decide the layout, or force a specific number of columns
) {
  
  # Guard: Check if spatial data was captured
  if (is.null(experiment_output$lists)) {
    stop("Spatial data is missing. You must run `run_segregation_model()` with `capture_spatial = TRUE`.")
  }
  
  # Check which runs actually exist in the data to prevent errors
  available_runs <- unique(experiment_output$lists$run_number)
  valid_runs <- intersect(runs_to_compare, available_runs)
  
  if (length(valid_runs) == 0) {
    stop("None of the requested runs exist in the experiment output.")
  }
  if (length(valid_runs) < length(runs_to_compare)) {
    warning("Some requested runs were not found and will be skipped.")
  }
  
  # Generate a list of plots using the existing plot_segregation_world function
  plot_list <- map(valid_runs, function(run_num) {
    
    # Generate the base plot
    p <- plot_segregation_world(
      experiment_output = experiment_output, 
      step_to_plot = step_to_plot, 
      run_to_plot = run_num
    )
    
    # Tweak the titles so they look cleaner side-by-side
    p <- p + 
      labs(title = paste("Run:", run_num)) + # Overwrite main title to save space
      theme(
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_blank() # Remove redundant subtitle
      )
    
    return(p)
  })
  
  # Combine them into a single side-by-side layout using patchwork
  combined_plot <- wrap_plots(plot_list, ncol = ncol) +
    plot_annotation(
      title = "Cross-Run Spatial Comparison",
      subtitle = paste("Step:", step_to_plot),
      theme = theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 12)
      )
    )
  
  return(combined_plot)
}
```


## animations

```{r}
library(tidyverse)
library(logolink)
library(colorspace)
library(magick)

animate_segregation_world <- function(
  experiment_output, 
  run_to_animate = 1, 
  fps = 10, 
  output_filename = "segregation_animation.gif"
) {
  
  if (is.null(experiment_output$lists)) {
    stop("Spatial data is missing. You must run `run_segregation_model()` with `capture_spatial = TRUE`.")
  }
  
  # 1. Isolate the specific run
  list_data <- experiment_output$lists |>
    filter(run_number == run_to_animate)
  
  if (nrow(list_data) == 0) {
    stop("No data found for run ", run_to_animate)
  }
  
  message("Pre-processing spatial data for animation...")
  
  # 2. Extract and parse all PATCHES across all steps
  all_patches <- list_data |>
    select(step, ends_with("_of_patches")) |>
    unnest(cols = ends_with("_of_patches")) |>
    rename_with(~ str_remove(.x, "_of_patches")) |>
    rename_with(~ str_replace_all(.x, "[\\?\\-]", "_")) |> 
    mutate(
      pcolor = parse_netlogo_color(as.numeric(pcolor)),
      pxcor = as.numeric(pxcor),
      pycor = as.numeric(pycor)
    )
  
  if (!"is_road_" %in% names(all_patches)) all_patches$is_road_ <- FALSE
  if (!"is_river_" %in% names(all_patches)) all_patches$is_river_ <- FALSE
  
  all_patches <- all_patches |>
    mutate(
      is_road_ = as.logical(is_road_),
      is_river_ = as.logical(is_river_),
      render_color = case_when(
        is_river_ ~ "#4682B4", 
        is_road_ ~ "#333333",  
        TRUE ~ pcolor
      )
    )
  
  # 3. Extract and parse all FAMILIES across all steps
  all_families <- list_data |>
    select(step, ends_with("_of_families")) |>
    unnest(cols = ends_with("_of_families")) |>
    rename_with(~ str_remove(.x, "_of_families")) |>
    mutate(
      color = parse_netlogo_color(as.numeric(color)),
      xcor = as.numeric(xcor),
      ycor = as.numeric(ycor)
    )
  
  # 4. Extract and parse all VENUES across all steps
  venue_cols <- names(list_data)[str_detect(names(list_data), "_of_venues$")]
  has_venues <- FALSE
  if (length(venue_cols) > 0) {
    all_venues <- list_data |>
      select(step, all_of(venue_cols)) |>
      unnest(cols = all_of(venue_cols))
    
    if (nrow(all_venues) > 0) {
      has_venues <- TRUE
      all_venues <- all_venues |>
        rename_with(~ str_remove(.x, "_of_venues")) |>
        mutate(
          base_color = parse_netlogo_color(as.numeric(color)),
          fill_color = colorspace::lighten(base_color, amount = 0.3),
          xcor = as.numeric(xcor),
          ycor = as.numeric(ycor)
        )
    }
  }

  # 5. Setup Temporary Directory for Frames
  frames_dir <- file.path(tempdir(), "netlogo_frames")
  if (!dir.exists(frames_dir)) dir.create(frames_dir)
  
  # Clear old frames if they exist from a previous run
  unlink(list.files(frames_dir, full.names = TRUE))
  
  steps <- sort(unique(list_data$step))
  frame_files <- character(length(steps))
  
  message(sprintf("Rendering %d frames. This may take a moment...", length(steps)))
  
  # 6. Render Loop
  for (i in seq_along(steps)) {
    current_step <- steps[i]
    
    # Subset data for this exact step
    step_patches <- all_patches |> filter(step == current_step)
    step_families <- all_families |> filter(step == current_step)
    
    p <- ggplot() +
      geom_raster(data = step_patches, aes(x = pxcor, y = pycor, fill = render_color)) +
      geom_point(data = step_families, aes(x = xcor, y = ycor, color = color), shape = 15, size = 2) +
      scale_fill_identity() + 
      scale_color_identity() + 
      coord_equal() +
      theme_void() +
      labs(
        title = "NetLogo World State",
        subtitle = sprintf("Run: %d | Step: %03d", run_to_animate, current_step)
      ) +
      theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, color = "gray40", family = "mono") # Mono prevents title jitter
      )
    
    if (has_venues) {
      step_venues <- all_venues |> filter(step == current_step)
      if (nrow(step_venues) > 0) {
        p <- p + geom_point(
          data = step_venues,
          aes(x = xcor, y = ycor, fill = fill_color),
          shape = 24, color = "black", size = 3.5, stroke = 0.8     
        )
      }
    }
    
    # Save the frame
    # Using %03d ensures frames are sorted correctly (e.g., frame_001.png)
    frame_name <- file.path(frames_dir, sprintf("frame_%03d.png", i))
    ggsave(filename = frame_name, plot = p, width = 6, height = 6, dpi = 150, bg = "white")
    frame_files[i] <- frame_name
  }
  
  message("Compiling GIF...")
  
  # 7. Compile and save the GIF using Magick
  animation <- magick::image_read(frame_files) |>
    magick::image_animate(fps = fps, dispose = "previous")
  
  magick::image_write(animation, output_filename)
  
  message(sprintf("Success! Animation saved to: %s", normalizePath(output_filename)))
  
  # Return the magick object so it can be viewed directly in the RStudio Viewer pane
  return(animation)
}
```


## trashold uttility (does not work. should not work)

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = c(0.1, 0.5, 0.9), # Now testing this!
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants, 
  repetitions = 5
)

# Render the new plot
print(new_experiment$plot)+
  guides(linetype = "none")
```

## low neighborhood preferences (beta) deceases segregation

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants2 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5, # Now testing this!
  "beta1-neighbors" = c(0.25, 0.75),
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.25,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment2 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants2, 
  repetitions = 5
)

# Render the new plot
print(new_experiment2$plot)
```

## low venue preferences (beta) increases segregation

gives an opportunities to move away from a home location segregation

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants3 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5, # Now testing this!
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = c(0.25, 0.75),
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment3 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants3, 
  repetitions = 5
)

# Render the new plot
print(new_experiment3$plot)
```

## Parks idk


```{r}
custom_constants4 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = c(0, 10),
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment4 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants4, 
  repetitions = 5
)

# Render the new plot
print(new_experiment4$plot)
```


## neighboorhood size (sanity check)

model cant converge with a really large distance?

```{r}
custom_constants5 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = c(1, 2, 10),
  "venues-number" = 5,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment5 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants5, 
  repetitions = 5
)

# Render the new plot
print(new_experiment5$plot)
```


## Number of venues

More venues --> less exposure (at least early)? everything else is same

```{r}
custom_constants6 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = c(0, 5, 25),
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment6 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants6, 
  repetitions = 5
)

# Render the new plot
print(new_experiment6$plot)
```

# road-scenario

roads increase exposure (compared to lack of roads), since they increase travel?

```{r}
custom_constants7 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 6,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty", "grid"),
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment7 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants7, 
  repetitions = 5
)

# Render the new plot
print(new_experiment7$plot)
```

## venue-exclusivity

lower exclusivity increases exposure

```{r}
custom_constants8 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 6,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = c(0, 0.5, 1),
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment8 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants8, 
  repetitions = 5
)

# Render the new plot
print(new_experiment8$plot)
```

## Interventions



### different road configurations

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants9 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("isolated-communities", "city-center-suburbia", "connected-communities"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment9 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants9, 
  repetitions = 5
)

# Render the new plot
print(experiment9$plot)
```

fully connected communities are hight on clustering, exposure and dissimilarity, compared to a isolated and city center 

### beta6-roads-group

```{r}
custom_constants11 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("isolated-communities"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0,
  "beta6-roads-group" = c(0, 0.5, 1),
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment11 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants11, 
  repetitions = 5
)

# Render the new plot
print(experiment11$plot)
```

### proportion-majority

it is hard to get exposure when your can't find any minorities


### intervention-scenario

#### central-park

central parks does nothing???

```{r}
custom_constants13 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 50,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "central-park"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment13 <- run_segregation_experiment(
  model_path = model_path, 
  constants = custom_constants13, 
  repetitions = 5,
  indices_to_plot = c("index_exposure", "index_dissimilarity", "index_concentration")
)

# Render the new plot
print(experiment13$plot)
```

#### mixed-grid

mixed grid of parks increases segregation, since there is no need to move and everything is around.

```{r}
custom_constants13 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 50,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "mixed-grid"),
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment13 <- run_segregation_model(
  capture_spatial = T,
  model_path = model_path, 
  constants = custom_constants13, 
  repetitions = 5,
  max_steps = 100
)

# Render the new plot
experiment13 |> 
  plot_segregation_results() +
  labs(caption = NULL)
  ggsave(width = 16, height = 9, filename = "expleriment_grid.png")

experiment13 |> plot_segregation_world(
  step_to_plot = 100, 
  run_to_plot = 7
) |> ggsave(filename = "expleriment_grid7.png")

experiment13 |> compare_segregation_runs(
  runs_to_compare = c(1, 2, 6, 7),
  step_to_plot = 100,
  ncol = 2 # Forces them into a single horizontal row
)

my_gif <- animate_segregation_world(
  experiment_output = experiment13,
  run_to_animate = 6,
  fps = 10,  # 10 frames per second means a 100-step run takes 10 seconds to watch
  output_filename = "run_2_mixed-grid.gif"
)
```


#### "majority-spread-minority-hubs"

"majority-spread-minority-hubs" there are not consequences when proportion is equal

```{r}
custom_constants13 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 50,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "majority-spread-minority-hubs"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment13 <- run_segregation_experiment(
  model_path = model_path, 
  constants = custom_constants13, 
  repetitions = 5
)

# Render the new plot
print(experiment13$plot)


custom_constants14 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 80,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "majority-spread-minority-hubs"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment14 <- run_segregation_experiment(
  model_path = model_path, 
  constants = custom_constants14, 
  repetitions = 5
)

# Render the new plot
print(experiment13$plot) 
print(experiment14$plot)
```

### "integrated-venues"

```{r}
custom_constants13 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 50,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "integrated-venues"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment13 <- run_segregation_model(
  capture_spatial = T,
  model_path = model_path, 
  constants = custom_constants13, 
  repetitions = 5,
  max_steps = 100
)



experiment13 |> 
  plot_segregation_results() +
  labs(caption = NULL) 
  ggsave(width = 16, height = 9, filename = "expleriment_integrated.png")

experiment13 |> plot_segregation_world(
  step_to_plot = 100, 
  run_to_plot = 8
) |> ggsave(filename = "expleriment_integrated8.png")

experiment13 |> compare_segregation_runs(
  runs_to_compare = c(1, 2, 6, 7),
  step_to_plot = 100,
  ncol = 2 # Forces them into a single horizontal row
)

my_gif <- animate_segregation_world(
  experiment_output = experiment13,
  run_to_animate = 8,
  fps = 10,  # 10 frames per second means a 100-step run takes 10 seconds to watch
  output_filename = "run_8_integrated.gif"
)

```

## Add world plots

```{r}
custom_constants13 <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 50,
  "neighborhood-distance" = 2,
  "venues-number" = 10,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty"),
  "intervention-scenario" = c("none", "integrated-venues"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

experiment13 <- run_segregation_model(
  capture_spatial = T,
    model_path = model_path, 
  constants = custom_constants13, 
  repetitions = 3,
  max_steps = 150
)

experiment13 |> 
  plot_segregation_results()

experiment13 |> plot_segregation_world(
  step_to_plot = 12, 
  run_to_plot = 6
)
```


### Animation, mostly for fun


```{r}
my_gif <- animate_segregation_world(
  experiment_output = experiment13,
  run_to_animate = 1,
  fps = 10,  # 10 frames per second means a 100-step run takes 10 seconds to watch
  output_filename = "run_1_segregation.gif"
)
```


### multiple worlds

```{r}
#| warning: false
library(tidyverse)



experiment13 |> compare_segregation_runs(
  runs_to_compare = c(1, 2, 4, 5),
  step_to_plot = 100,
  ncol = 2 # Forces them into a single horizontal row
)
```


## continious 

```{r}
library(tidyverse)
library(logolink)
library(geomtextpath)

# Update this to point to your saved NetLogo file
model_path <- file.path("netlogo", "roads", "Segregation Interventions.nlogox")
model_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Interventions.nlogox"

# Define a consistent baseline environment
base_config <- list(
  "density-percent" = 70,
  "number-of-ethnicities" = 2,
  "proportion-majority" = 60,
  "neighborhood-distance" = 2,
  "venues-number" = 0,
  "venue-catchment-distance" = 7, 
  "venue-exclusivity" = 1,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "beta5-park-group" = 0.5,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "road-scenario" = "grid",
  "road-density" = 25,
  "road-connectiveness" = 80,
  "parks-number" = 0 # Disabled to isolate the effect of venues
)
```

### integration structured 0-100

```{r}
exp1_integration <- run_segregation_model(
  model_path = model_path,
  repetitions = 5,
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = c(0, 25, 50, 75, 100),
      `intervention-scenario` = "structured",
      `venues-number` = 10
    )
)

plot_segregation_results(exp1_integration, c("index_concentration", "index_dissimilarity"))
```


### random vs structured

```{r}
exp2_scenario <- run_segregation_model(
  model_path = model_path,
  repetitions = 5,
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = 100,
      `intervention-scenario` = c("random", "structured"),
      `venues-number` = 10
    )
)

plot_segregation_results(exp2_scenario, c("index_concentration", "index_dissimilarity"))
```


### structured venues 0 - 20

```{r}
exp3_volume <- run_segregation_model(
  model_path = model_path,
  repetitions = 5,
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = 100,
      `intervention-scenario` = "structured",
      `venues-number` = c(0, 5, 10, 20)
    )
)

plot_segregation_results(exp3_volume, c("index_concentration", "index_dissimilarity"))
```


### all


```{r}
# Run the full factorial design
exp_full_space <- run_segregation_model(
  model_path = model_path,
  repetitions = 3, # Lowered repetitions to manage execution time
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = c(0, 50, 100),
      `intervention-scenario` = c("random", "structured"),
      `venues-number` = c(0, 5, 10, 15)
    )
)

# Custom visualization for the full space
exp_full_space$data |> 
  filter(step <= exp_full_space$max_steps) |> 
  # Isolate only the columns we need to prevent memory bloat
  select(
    run_number, step, 
    index_concentration, index_dissimilarity, 
    proportion_of_integration, intervention_scenario, venues_number
  ) |> 
  pivot_longer(
    cols = c(index_concentration, index_dissimilarity),
    names_to = "index", 
    values_to = "value"
  ) |> 
  mutate(
    index = str_remove(index, "index_"),
    # Create a unique group id for each run and metric type to prevent line-bridging
    grouping_id = str_c(run_number, index)
  ) |> 
  ggplot(aes(
    x = step, 
    y = value, 
    color = factor(proportion_of_integration), 
    group = grouping_id
  )) +
  geom_line(aes(linetype = index), alpha = 0.3, linewidth = 1) +
  # Tidy principle: Create a trellis graph to map multiple dimensions clearly
  facet_grid(
    intervention_scenario ~ venues_number, 
    labeller = labeller(venues_number = label_both)
  ) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    title = "Segregation Dynamics Across the Full Intervention Design Space",
    subtitle = "Rows: Placement Strategy | Columns: Total Venues | Color: Proportion Integrated (%)",
    x = "Simulation Step",
    y = "Massey & Denton Index Value",
    color = "Integration (%)",
    linetype = "Metric"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "gray95", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```

```{r}
# 1. Run the full factorial design with 100 repetitions
exp_full_space_100 <- run_segregation_model(
  model_path = model_path,
  repetitions = 100, 
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = c(0, 50, 100),
      `intervention-scenario` = c("random", "structured"),
      `venues-number` = c(0, 5, 10, 15)
    )
)

# 2. Extract and tidy the base data for individual runs
plot_data <- exp_full_space_100$data |> 
  filter(step <= exp_full_space_100$max_steps) |> 
  select(
    run_number, step, 
    index_concentration, index_dissimilarity, 
    proportion_of_integration, intervention_scenario, venues_number
  ) |> 
  pivot_longer(
    cols = c(index_concentration, index_dissimilarity),
    names_to = "index", 
    values_to = "value"
  ) |> 
  mutate(
    index = str_remove(index, "index_"),
    # grouping_id ensures geom_line doesn't bridge different runs or metrics together
    grouping_id = str_c(run_number, index) 
  )

# 3. Calculate the mean values per condition and step
mean_data <- plot_data |> 
  group_by(step, index, proportion_of_integration, intervention_scenario, venues_number) |> 
  summarize(mean_value = mean(value, na.rm = TRUE), .groups = "drop")

# 4. Build the layered plot
p <- ggplot() +
  # Layer 1: Individual runs (highly transparent to show density/variance)
  geom_line(
    data = plot_data,
    aes(
      x = step, 
      y = value, 
      color = factor(proportion_of_integration), 
      group = grouping_id,
      linetype = index
    ),
    alpha = 0.05
  ) +
  # Layer 2: Condition Means (fully opaque to anchor the trend)
  geom_line(
    data = mean_data,
    aes(
      x = step, 
      y = mean_value, 
      color = factor(proportion_of_integration),
      linetype = index
    ),
    alpha = 1,
    linewidth = 1.2
  ) +
  facet_grid(
    intervention_scenario ~ venues_number, 
    labeller = labeller(venues_number = label_both)
  ) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    title = "Segregation Dynamics: 100 Repetitions per Condition",
    subtitle = "Rows: Placement Strategy | Columns: Total Venues | Color: Proportion Integrated (%)",
    x = "Simulation Step",
    y = "Massey & Denton Index Value",
    color = "Integration (%)",
    linetype = "Metric"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "gray95", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )

# 5. Save the plot to disk
ggsave(
  filename = "segregation_design_space_100reps.png", 
  plot = p, 
  width = 14,    # Wide enough to accommodate the 4 columns
  height = 8, 
  dpi = 300, 
  bg = "white"   # Ensures background isn't transparent in some viewers
)
```

```{r}
metrics_to_plot <- c("index_exposure", "index_dissimilarity", "index_clustering", "index_concentration")

# ==============================================================================
# EXPERIMENT 1: VENUE PLACEMENT & INTEGRATION
# ==============================================================================

# 1. Run the simulation
exp_venues_100 <- run_segregation_model(
  model_path = model_path,
  repetitions = 100, 
  max_steps = 125,
  constants = base_config |> 
    list_modify(
      `proportion-of-integration` = c(0, 25, 50, 75, 100),
      `intervention-scenario` = c("random", "structured"),
      `venues-number` = c(0, 3, 9, 18)
    )
)

# 2. Save pure simulation data to CSV
write_csv(exp_venues_100$data, "venue_experiments_100reps.csv")

# 3. Tidy the data for plotting
venue_data_long <- exp_venues_100$data |> 
  filter(step <= exp_venues_100$max_steps) |> 
  select(
    run_number, step, 
    all_of(metrics_to_plot), 
    proportion_of_integration, intervention_scenario, venues_number
  ) |> 
  pivot_longer(
    cols = starts_with("index_"),
    names_to = "index", 
    values_to = "value"
  )

# 4. Iterate and plot each metric separately
walk(metrics_to_plot, \(current_metric) {
  
  plot_data <- venue_data_long |> filter(index == current_metric)
  
  mean_data <- plot_data |> 
    group_by(step, proportion_of_integration, intervention_scenario, venues_number) |> 
    summarize(mean_value = mean(value, na.rm = TRUE), .groups = "drop")
  
  metric_name <- str_to_title(str_replace(current_metric, "index_", ""))
  
  p <- ggplot() +
    # Layer 1: Individual variance (solid lines, low alpha)
    geom_line(
      data = plot_data,
      aes(
        x = step, 
        y = value, 
        color = factor(proportion_of_integration), 
        group = run_number
      ),
      alpha = 0.05,
      linetype = "solid"
    ) +
    # Layer 2: Mean trend with embedded textpath
    geomtextpath::geom_textpath(
      data = mean_data,
      aes(
        x = step, 
        y = mean_value, 
        color = factor(proportion_of_integration),
        label = paste0(proportion_of_integration, "%")
      ),
      offset = unit(5, "pt"),
      gap = TRUE,
      text_smoothing = 95,
      alpha = 1,
      linewidth = 1.2,
      linetype = "solid"
    ) +
    facet_grid(
      intervention_scenario ~ venues_number, 
      labeller = labeller(venues_number = label_both)
    ) +
    scale_color_viridis_d(option = "plasma", end = 0.8) +
    labs(
      title = paste(metric_name, "Dynamics (Venue Interventions)"),
      subtitle = "Rows: Strategy | Columns: Total Venues | Labeled by Integration %",
      x = "Simulation Step",
      y = metric_name
    ) +
    theme_minimal() +
    theme(
      legend.position = "none", # Legend removed since lines are directly labeled
      strip.background = element_rect(fill = "gray95", color = NA),
      strip.text = element_text(face = "bold"),
      panel.spacing = unit(1, "lines")
    )
  
  ggsave(
    filename = paste0("plot_venues_", current_metric, ".png"), 
    plot = p, 
    width = 12, height = 8, dpi = 300, bg = "white"
  )
})


# ==============================================================================
# EXPERIMENT 2: ROAD INFRASTRUCTURE
# ==============================================================================

# 1. Run the road simulation (Isolating roads by turning venues off)
exp_roads_100 <- run_segregation_model(
  model_path = model_path,
  repetitions = 100, 
  max_steps = 100,
  constants = base_config |> 
    list_modify(
      `road-density` = c(25, 50, 100),
      `road-connectiveness` = c(50, 75, 100),
      `venues-number` = 0 
    )
)

# 2. Save road simulation data to CSV
write_csv(exp_roads_100$data, "road_experiments_100reps.csv")

# 3. Tidy the data for plotting
road_data_long <- exp_roads_100$data |> 
  filter(step <= exp_roads_100$max_steps) |> 
  select(
    run_number, step, 
    all_of(metrics_to_plot), 
    road_density, road_connectiveness
  ) |> 
  pivot_longer(
    cols = starts_with("index_"),
    names_to = "index", 
    values_to = "value"
  )

# 4. Iterate and plot each metric separately
walk(metrics_to_plot, \(current_metric) {
  
  plot_data <- road_data_long |> filter(index == current_metric)
  
  mean_data <- plot_data |> 
    group_by(step, road_density, road_connectiveness) |> 
    summarize(mean_value = mean(value, na.rm = TRUE), .groups = "drop")
  
  metric_name <- str_to_title(str_replace(current_metric, "index_", ""))
  
  p <- ggplot() +
    geom_line(
      data = plot_data,
      aes(
        x = step, 
        y = value, 
        color = factor(road_density), 
        group = run_number
      ),
      alpha = 0.05,
      linetype = "solid"
    ) +
    geomtextpath::geom_textpath(
      data = mean_data,
      aes(
        x = step, 
        y = mean_value, 
        color = factor(road_density),
        label = paste0("Density: ", road_density)
      ),
      offset = unit(5, "pt"),
      gap = TRUE,
      text_smoothing = 95,
      alpha = 1,
      linewidth = 1.2,
      linetype = "solid"
    ) +
    facet_wrap(~ road_connectiveness, labeller = labeller(road_connectiveness = label_both)) +
    scale_color_viridis_d(option = "viridis", end = 0.8) +
    labs(
      title = paste(metric_name, "Dynamics (Road Infrastructure)"),
      subtitle = "Panels: Road Connectiveness | Labeled by Road Density",
      x = "Simulation Step",
      y = metric_name
    ) +
    theme_minimal() +
    theme(
      legend.position = "none",
      strip.background = element_rect(fill = "gray95", color = NA),
      strip.text = element_text(face = "bold"),
      panel.spacing = unit(1, "lines")
    )
  
  ggsave(
    filename = paste0("plot_roads_", current_metric, ".png"), 
    plot = p, 
    width = 12, height = 6, dpi = 300, bg = "white"
  )
})
```

