---
title: "experiments"
format: html
---

```{r}
library(tidyverse)
library(logolink)
```

```{r}
model_path <- file.path("netlogo", "roads", "Segregation Utility.nlogox")
model_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"
```

```{r}
setup_file <- create_experiment(
  name = "Venue Segragation Model Analysis",
  repetitions = 3,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 200,
  metrics = c(
  'index-exposure',
  'index-dissimilarity',
  'index-clustering',
  'index-centrality',
  'index-concentration'
  ),
  constants = list(
    "density-percent" = c(70),
    "%-similar-wanted" = 30,
    "number-of-ethnicities" = 2,
    "%-different-wanted" = 0,
    "neighborhood-distance" = c(2),
    "venues-number" = c(5),
    "venue-catchment-distance" = c(1, 3, 10, 15),
    "venue-exclusivity" = 1,
    "utility-threshold" = 0.1,
    "beta1-neighbors" = c(0.5),
    "beta2-venue-group" = 0.5,
    "beta3-venue-loc" = 0,
    "beta4-home-venues" = 0,
    "road-scenario" = 'empty',
    "parks-number" = 3,
    "park-catchment-distance" = 2,
    "beta5-park-group" = 0.25,
    "travel-distance" = 5,
    "road-multiplier" = 1,
    "real-travel?" = F,
    "random-seed?" = F
  )
)


```

```{r}
results <-
  model_path |>
  run_experiment(
    setup_file = setup_file,
    output = c("table")
  )
```

```{r}
library(magrittr)
data <-
  results |>
  extract2("table") 
```


### Segregation Indices Over Time by Neighborhood Distance

```{r}
data |> 
  filter(step <= 100) |>  
  select(
    run_number,
    step,
    starts_with("index_"),
    venue_catchment_distance         
  ) |> 
  pivot_longer(
    cols = starts_with("index_"),
    names_to = "index",
    values_to = "value"
  ) |> 
  filter(index %in% c("index_exposure", "index_dissimilarity", 
                      "index_clustering")) |>
  filter(venue_catchment_distance %in% c(1, 15)) |>
  mutate(grouping_combo_run = str_c(run_number, index, venue_catchment_distance),
         grouping_combo_condition = str_c(index, venue_catchment_distance)) |> 
  ggplot(aes(x = step, y = value, color = venue_catchment_distance |> factor(), group = grouping_combo_run, linetype = index)) +
  geom_line(alpha = 0.25)+
  # facet_wrap(~ venue_catchment_distance)+
  geomtextpath::geom_textpath(
    aes(group = index, label = index), 
    offset = unit(5, "pt"), gap = T, text_smoothing = 90)+
  labs(title = "Segregation Indices Over Time (venue_catchment_distance)", x = "Step", 
       y = "Index Value", color = "venue_catchment_distance") +
  theme_minimal()+
  theme(legend.position = "bottom")
       
```


### function

```{r}

input_to_vis <- function(model_path, constrain, values = c()){
  # function takes the model path
  # the name of the constraint to be varied and a vector of values for that constraint
  # it creates an experiment file and runs the experiment, returning the results as a graph
  
}
```



```{r}
model_path <-
  find_netlogo_home() |>
  file.path(
    "models",
    "IABM Textbook",
    "chapter 4",
    "Wolf Sheep Simple 5.nlogox"
  )

setup_file <- create_experiment(
  name = "Wolf Sheep Simple Model Analysis",
  setup = "setup",
  go = "go",
  time_limit = 500,
  run_metrics_condition = 'ticks mod 100 = 0',
  metrics = c(
    '[xcor] of sheep',
    '[ycor] of sheep',
    '[xcor] of wolves',
    '[ycor] of wolves',
    '[pxcor] of patches',
    '[pycor] of patches',
    '[pcolor] of patches'
  ),
  constants = list(
    "number-of-sheep" = 100,
    "number-of-wolves" = 15,
    "movement-cost" = 0.5,
    "grass-regrowth-rate" = 0.3,
    "energy-gain-from-grass" = 2,
    "energy-gain-from-sheep" = 5
  )
)
```

```{r}
create_experiment_template(model_path, insert_at_cursor = TRUE)

setup_file <- create_experiment(
  name = "Auto-Generated Analysis",
  repetitions = 1,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 200,
  metrics = c(
    "percent-similar",
    "percent-unhappy",
    "index-exposure",
    "index-dissimilarity",
    "index-clustering",
    "index-centrality",
    "index-concentration",
    "[pxcor] of patches",
    "[pycor] of patches",
    "[pcolor] of patches"
  ),
  constants = list(
    "density-percent" = c(71.0),
    "%-similar-wanted" = c(30.0),
    "number-of-ethnicities" = c(2.0),
    "%-different-wanted" = c(0.0),
    "neighborhood-distance" = c(1.0),
    "venues-number" = c(4.0),
    "venue-catchment-distance" = c(11.0),
    "venue-exclusivity" = c(1.0),
    "utility-threshold" = c(0.1),
    "beta1-neighbors" = c(0.1),
    "beta2-venue-group" = c(0.1),
    "beta3-venue-loc" = c(0.1),
    "beta4-home-venues" = c(0.1),
    "parks-number" = c(1.0),
    "park-catchment-distance" = c(10.0),
    "beta5-park-group" = c(0.1),
    "travel-distance" = c(9.0),
    "road-multiplier" = c(1.0),
    "real-travel?" = c(FALSE),
    "random-seed?" = c(FALSE),
    "new-venue-color" = c(15.0),
    "visual-mode" = c("abstract"),
    "road-scenario" = c("empty")
  )
)
```

### funtion 

```{r}
library(tidyverse)
library(logolink)
library(geomtextpath)

run_segregation_experiment <- function(
  model_path,
  repetitions = 3,
  max_steps = 100,
  indices_to_plot = c("index_exposure", "index_dissimilarity", "index_clustering"),
  # Pass constants as a list so you can easily modify them per run
  constants = list(
    "density-percent" = 70,
    "%-similar-wanted" = 30,
    "number-of-ethnicities" = 2,
    "%-different-wanted" = 0,
    "neighborhood-distance" = 2,
    "venues-number" = 5,
<<<<<<< HEAD
    "venue-catchment-distance" = 7, # Example of a tested/varying parameter
=======
    "venue-catchment-distance" = c(1, 15), # Example of a tested/varying parameter
>>>>>>> 3347dcd28db32ab73a9fc8f6442549182fb0bcfc
    "venue-exclusivity" = 1,
    "utility-threshold" = 0.1,
    "beta1-neighbors" = 0.5,
    "beta2-venue-group" = 0.5,
    "beta3-venue-loc" = 0,
    "beta4-home-venues" = 0,
    "road-scenario" = "empty",
    "parks-number" = 3,
    "park-catchment-distance" = 2,
    "beta5-park-group" = 0.25,
    "travel-distance" = 5,
    "road-multiplier" = 1,
    "real-travel?" = FALSE,
    "random-seed?" = FALSE
  )
) {
  
  # 1. Define the experiment setup
  setup_file <- create_experiment(
    name = "Venue Segregation Model Analysis",
    repetitions = repetitions,
    run_metrics_every_step = TRUE,
    setup = "setup",
    go = "go",
    time_limit = max_steps + 50, # Buffer to ensure we hit max_steps safely
    metrics = c(
      'index-exposure',
      'index-dissimilarity',
      'index-clustering',
      'index-centrality',
      'index-concentration'
    ),
    constants = constants
  )
  
  # 2. Run the experiment
  results <- model_path |>
    run_experiment(
      setup_file = setup_file,
      output = c("table")
    )
  
  # 3. Extract the table natively
  data <- results[["table"]]
  
  # 4. Identify the varying parameter to dynamically color the plot
  # Finds any constant with length > 1 (e.g., c(1, 15))
  varying_params <- names(constants)[lengths(constants) > 1]
  
  if (length(varying_params) == 0) {
    legend_title <- "run_number"
    varying_col <- "run_number"
  } else {
    legend_title <- varying_params[1]
    # NetLogo hyphenated params become underscored cols in R
    varying_col <- str_replace_all(legend_title, "-", "_") 
  }
  
  # 5. Build dynamic title string from all constants
  param_strings <- imap_chr(constants, ~ paste0(.y, ": ", paste(.x, collapse = ", ")))
  dynamic_subtitle <- str_wrap(paste(param_strings, collapse = " | "), width = 110)
  
  # 6. Process data and generate the plot
  p <- data |> 
    filter(step <= max_steps) |> 
    select(
      run_number,
      step,
      starts_with("index_"),
      all_of(varying_col) 
    ) |> 
    pivot_longer(
      cols = starts_with("index_"),
      names_to = "index",
      values_to = "value"
    ) |> 
    filter(index %in% indices_to_plot) |>
    mutate(
      # Tidy evaluation: Use .data[[varying_col]] to reference the column dynamically
      grouping_combo_run = str_c(run_number, index, .data[[varying_col]]),
      # Clean up the index names for prettier labels
      index_label = str_remove(index, "index_") 
    ) |> 
    ggplot(aes(
      x = step, 
      y = value, 
      color = factor(.data[[varying_col]]), 
      group = grouping_combo_run, 
      linetype = index_label
    )) +
    geom_line(alpha = 0.25) +
    geomtextpath::geom_textpath(
      aes(group = index_label, label = index_label), 
      offset = unit(5, "pt"), 
      gap = TRUE, 
      text_smoothing = 90,
      color = "gray20"
    ) +
    labs(
      title = "Segregation Indices Over Time",
      subtitle = varying_col, # Show the name of the varying parameter in the subtitle
      caption  = dynamic_subtitle,
      x = "Step", 
      y = "Index Value", 
      color = legend_title,
      linetype = "Index"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      plot.subtitle = element_text(size = 11, color = "gray20", margin = margin(b = 10)),
      plot.caption = element_text(size = 7, color = "gray50", hjust = 0, margin = margin(t = 15))
    )+
    guides(linetype = "none")
  
  # Return both the raw data and the plot as a list so you can access either
  return(list(
    data = data,
    plot = p
  ))
}
```

## trashold uttility (does not work. should not work)

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = c(0.1, 0.5, 0.9), # Now testing this!
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
<<<<<<< HEAD
  "beta5-park-group" = 0,
  "beta6-roads-group" = 0.5,
=======
  "beta5-park-group" = 0.25,
>>>>>>> 3347dcd28db32ab73a9fc8f6442549182fb0bcfc
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants, 
  repetitions = 5
)

# Render the new plot
print(new_experiment$plot)+
  guides(linetype = "none")
```

## low neighborhood preferences (beta) deceases segregation

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants2 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5, # Now testing this!
  "beta1-neighbors" = c(0.25, 0.75),
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.25,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment2 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants2, 
  repetitions = 5
)

# Render the new plot
print(new_experiment2$plot)
```

## low venue preferences (beta) increases segregation

gives an opportunities to move away from a home location segregation

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants3 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5, # Now testing this!
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = c(0.25, 0.75),
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment3 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants3, 
  repetitions = 5
)

# Render the new plot
print(new_experiment3$plot)
```

## Parks idk


```{r}
custom_constants4 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = c(0, 10),
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment4 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants4, 
  repetitions = 5
)

# Render the new plot
print(new_experiment4$plot)
```


## neighboorhood size (sanity check)

model cant converge with a really large distance?

```{r}
custom_constants5 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = c(1, 2, 10),
  "venues-number" = 5,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment5 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants5, 
  repetitions = 5
)

# Render the new plot
print(new_experiment5$plot)
```


## Number of venues

More venues --> less exposure (at least early)? everything else is same

```{r}
custom_constants6 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = c(0, 5, 25),
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment6 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants6, 
  repetitions = 5
)

# Render the new plot
print(new_experiment6$plot)
```

# road-scenario

roads increase exposure (compared to lack of roads), since they increase travel?

```{r}
custom_constants7 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 6,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("empty", "grid"),
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment7 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants7, 
  repetitions = 5
)

# Render the new plot
print(new_experiment7$plot)
```

## venue-exclusivity

lower exclusivity increases exposure

```{r}
custom_constants8 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 6,
  "venue-catchment-distance" = 10, 
  "venue-exclusivity" = c(0, 0.5, 1),
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = "empty",
  "parks-number" = 0,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
new_experiment8 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants8, 
  repetitions = 5
)

# Render the new plot
print(new_experiment8$plot)
```

<<<<<<< HEAD
## Interventions



### different road configurations

```{r}
my_path <- "C:/Users/vsesu60/OneDrive - Linköpings universitet/Documents/R/vis-soc/netlogo/roads/Segregation Utility.nlogox"

custom_constants9 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("isolated-communities", "city-center-suburbia", "connected-communities"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0,
  "beta6-roads-group" = 0.5,
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment9 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants9, 
  repetitions = 5
)

# Render the new plot
print(experiment9$plot)
```

fully connected communities are hight on clustering, exposure and dissimilarity, compared to a isolated and city center 

### beta6-roads-group

```{r}
custom_constants11 <- list(
  "density-percent" = 70,
  "%-similar-wanted" = 30,
  "number-of-ethnicities" = 2,
  "%-different-wanted" = 0,
  "neighborhood-distance" = 2,
  "venues-number" = 5,
  "venue-catchment-distance" = 10, # Keep this static this time
  "venue-exclusivity" = 1,
  "utility-threshold" = 0.5,
  "beta1-neighbors" = 0.5,
  "beta2-venue-group" = 0.5,
  "beta3-venue-loc" = 0,
  "beta4-home-venues" = 0,
  "road-scenario" = c("isolated-communities"),
  "parks-number" = 3,
  "park-catchment-distance" = 2,
  "beta5-park-group" = 0,
  "beta6-roads-group" = c(0, 0.5, 1),
  "travel-distance" = 5,
  "road-multiplier" = 1,
  "real-travel?" = FALSE,
  "random-seed?" = FALSE
)

# Run with custom parameters
experiment11 <- run_segregation_experiment(
  model_path = my_path, 
  constants = custom_constants11, 
  repetitions = 5
)

# Render the new plot
print(experiment11$plot)
```

=======
>>>>>>> 3347dcd28db32ab73a9fc8f6442549182fb0bcfc
